Relationships in the extremes and their influence on competition and coexistence: Supporting information

### 2022-01-26 paper/SupMat.Rnw
edited SI section 9 and a few other minor changes

\DeclareMathOperator{\EX}{E}% expected value
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\VarX}{var}
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\CovX}{cov}
\DeclareMathOperator{\cov}{cov}
\DeclareMathOperator{\mean}{mean}
\DeclareMathOperator{\se}{se}
\DeclareMathOperator{\sd}{sd}
\DeclareMathOperator{\Prob}{P}

@@

The decomposition of storage effects and quantification of contributions of ATAs depends on 
the removal of asymmetric tail associations from the bivariate random variable
$(E_i,C_{i \bs i})$, while maintaining its correlation and marginal distributions, to produce a new
random variable $(E_i^\nat,C_{i \bs i}^\nat)$.
In the Theory section of the main text, we decribed this removal process
when the distributions of $E_i$ and $C_{i \bs i})$ are explicitly known, but for the 
diatom example explored for goal G4 of the main text (see also SI section \ref{sect:diatom_model}), 
we only have samples from $(E_i,C_{i \bs i})$, based on simulations. We here describe how to
produce a sample from $(E_i^\nat,C_{i \bs i}^\nat)$ in this modified context.

In the paper section \ref{MT-section:theory}, we describe analytical methods for generating the partial sharp variables, $E_i^\nat$ and $C_{i \bs i}^\nat$, which are distributed in the same way as the defined $E_i$ and $C_{i \bs i}$, but are symmetric in the tails and have a normal copula structure.
The protocol for samples of coexistence drivers in empirical applications is as follows.
Let $\olE_i$ be a sample of a environmental fluctuation and $\olC_{i \bs i}$ be a sample of competitive pressure. Let $\Theta_{\olE}$ and $\Theta_{\olC}$ be standard normal distributions obtained by normalizing the ranks of $\olE_i$ and $\olC_{i \bs i}$, 
such that $\Theta_{\olE} = \frac{rank(\olE_i)}{n_{\olE_i}+1} \circ \varphi^{-1}$ and $\Theta_{\olC} = \frac{rank(\olC_{i \bs i})}{n_{\olC_{i \bs i}}+1} \circ \varphi^{-1}$, where $\varphi$ is the CDF of a standard normal distribution. The normalized correlation of $\olE_i$ and $\olC_{i \bs i}$, denoted by $r$, is the covariance of $\Theta_{\olE}$ and $\Theta_{\olC}$.  
We then simulate an $S$ number of bivariate normal distributions $N_s\left(\vec{0}, \Sigma \right)$
with covariance matrix $\Sigma = \begin{pmatrix} 1&\ddot{\rho} \\ \ddot{\rho}&1\end{pmatrix}$. 
The columns of $(\olE_{i}, \olC_{i \bs i})$ are permutated and saved $S$ amount of times, 
such that the rank of each column matches with the rank of each column of $N_s$. 
To obtain $(\olE_{i}^\nat, \olC_{i \bs i}^\nat)$, the median of the permuations is computed. 
Here, $\ddot\rho$ can be chosen a number of ways such that the Pearson, Spearman, or Kendall 
correlation of $(\olE_{i}^\nat, \olC_{i \bs i}^\nat)$ matches that of $(\olE_{i}, \olC_{i \bs i})$
or such that $\ddot\rho = r$. The code for this protocol is in \url{"diatom/partialSharp_fxns.R"} and was used in our application to the diatom chemostat system.

>>>>

\DeclareMathOperator{\EX}{E}% expected value
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\VarX}{var}
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\CovX}{cov}
\DeclareMathOperator{\cov}{cov}
\DeclareMathOperator{\mean}{mean}
\DeclareMathOperator{\se}{se}
\DeclareMathOperator{\sd}{sd}
\DeclareMathOperator{\Prob}{P}
\DeclareMathOperator{\rk}{rank}
\DeclareMathOperator{\nrk}{nrank}

@@

The decomposition of storage effects and quantification of contributions of ATAs depends on 
removing asymmetric tail associations from the bivariate random variable
$(E_i,C_{i \bs i})$, while maintaining its correlation and marginal distributions, to produce a new
random variable $(E_i^\nat,C_{i \bs i}^\nat)$.
In the Theory section of the main text, we decribed this removal process
when the distributions of $E_i$ and $C_{i \bs i}$ are known exactly, but for the 
diatom example explored for goal G4 of the main text (see also SI section \ref{sect:diatom_model}), 
we only had samples from $(E_i,C_{i \bs i})$, based on simulations. We here describe how to
produce a sample from $(E_i^\nat,C_{i \bs i}^\nat)$ in this modified context. 
Let $(E_{i,k},C_{i \bs i,k})$, $k=1,\ldots,T$ denote the samples from $(E_i,C_{i \bs i})$.
Let $\rk(E_{i,k})$ be the rank of $E_{i,k}$ in the set 
of $E_{i,k}$, $k=1,\ldots,T$, where the smallest value has rank $1$. Likewise 
define $\rk(C_{i \bs i,k})$ as the rank of $C_{i \bs i,k}$ in the set
of $C_{i \bs i,k}$, $k=1,\ldots,T$. Define the ``normalized rank''
$\nrk(E_{i,k})$ as $\rk(E_{i,k})/(T+1)$, and likewise define 
$\nrk(C_{i \bs i,k})$ as $\rk(C_{i \bs i,k})/(T+1)$. Letting $\varphi$ be the CDF
of a standard normal distribution, the samples $\varphi^{-1}(\nrk(E_{i,k}))$, $k=1,\ldots,T$
are standard normally distributed, as are the samples 
$\varphi^{-1}(\nrk(C_{i \bs i,k}))$, $k=1,\ldots,T$. Let $\rho_i$ be the covariance 
of these samples. We then define a bivariate normal random variable 
$(e_i,c_{i \bs i})$ such that $\E(e_i)=\E(c_{i \bs i})=0$, $\var(e_i)=\var(c_{i \bs i})=1$,
and we denote $\cov(e_i,c_{i \bs i})$ by $\ddot{\rho}_i$. We used $\ddot{\rho}_i = \rho_i$.
Generating a sample $(e_{i,k},c_{i \bs i,k})$, $k=1,\ldots,T$ from this distribution,
we permute the ordered set $(E_{i,1},E_{i,2},\ldots,E_{i,T})$ so that the ranks of the 
permuted ordered set equal the ranks of the corresponding elements of the ordered
set $(e_{i,1},e_{i,2},\ldots,e_{i,T})$. We likewise permute the ordered set 
$(C_{i \bs i,1},C_{i \bs i,2},\ldots,C_{i \bs i,T})$ so the ranks of the permuted result
match the ranks of $(c_{i \bs i,1},c_{i \bs i,2},\ldots,c_{i \bs i,T})$.
This step works as long as there are no ties among the $E_{i,k}$ or among
the $C_{i \bs i,k}$. There were no ties when we applied these methods to the diatom system. 
The two permuted ordered sets, here denoted $(E_{i,1}^\nat,E_{i,2}^\nat,\ldots,E_{i,T}^\nat)$
and $(C_{i \bs i,1}^\nat,C_{i \bs i,2}^\nat,\ldots,C_{i \bs i,T}^\nat)$,
when combine to make ordered pairs, form the desired sample from 
$(E_i^\nat,C_{i \bs i}^\nat)$. Because $(E_{i,1}^\nat,\ldots,E_{i,T}^\nat)$
is a permutation of $(E_{i,1},\ldots,E_{i,T})$ and 
$(C_{i \bs i,1}^\nat,\ldots,C_{i \bs i,T}^\nat)$ is a permutation of
$(C_{i \bs i,1},\ldots,C_{i \bs i,T})$, marginal distributions of our
sample from $(E_i^\nat,C_{i \bs i}^\nat)$ were the same as those of our original 
sample from $(E_i, C_{i \bs i})$. The choice of bivariate normal
$(e_i,c_{i \bs i})$ ensures that our resulting sample shows symmetric tail association.
Finally, we caution that sampling variation means any quantity computed from our resulting sample 
$(E_{i,k}^\nat,C_{i \bs i,k}^\nat)$, $k=1,\ldots,T$ should be recomputed for 
multiple such samples, all computed via the same process, and a median of 
the resulting distribution of values should be used.
